Estimating duration in the presence of misclassification

### 2024-02-22 introduction.tex
start outling intro

\begin{itemize}
    \item Disease burden of pandemic and ongoing burden of COVID
    \item To mitigate its effects, need to understand incidence and transmission but cannot observe directly
    \item Importance of surveillance objectives and concepts: incidence, transmission, basic reproduction number, effective reproduction number
    \item Expected future pandemics, methods should be reusable
\end{itemize}

% - Natural history of SARS-CoV-2 infection
% 	- Stylised graphic of phases: infection, pre-infection positive, infectious positive, post-infectious positive, negative
% 	- How it spreads
% - Why is incidence and prevalence estimation challenging?
% 	- Asymptomatic infection
% 	- Ascertainment bias
% - How do we estimate prevalence
% 	- CIS and REACT
% - Issues with current incidence estimation

\section{SARS-CoV-2 real-time surveillance}

\begin{itemize}
    \item Most commonly used types of data: severe events (deaths/hospitalisations), widespread testing (pillar 2), 
    \item Advantages of designed surveys (REACT, CIS)
    \item Use of cohorts (ATACCC)
    \item Statistical modelling performed during the pandemic. Discuss broadly but especially focus on those using CIS or REACT. Discuss my contribution in this context.
\end{itemize}

\subsection{Modelling}

\subsection{Surveillance data}

@@

\section{Historical context}

@@

\subsection{Surveillance}

\begin{itemize}
    \item Reproduction numbers
    \item Incidence: almost never observed directly
\end{itemize}

\subsection{Backcalculation}

@@

\subsection{Mechanistic models}

@@

\subsection{Duration} \label{inc-prev:sec:duration}

@@

@@

@@

The ATACCC data used is found in the same location.
CIS data can only be accessed by accredited researchers within the Office for National Statistic's Secure Research Service\todo{insert details of applying to SRS}.
Where possible, aggregated, non-disclosive versions of the dataset are available alongside the code.

>>>>

COVID-19 (Coronavirus Disease 2019), the disease caused by the virus SARS-CoV-2 (Severe Acute Respiratory Syndrome Coronavirus 2), is estimated to have killed around\todo{insert excess deaths estimates by Economist and/or WHO}.
While it is tempting to consider this a rare or extreme event, history suggests otherwise (see \cref{intro:fig:pandemic-timeline}).
\begin{figure}
    \includegraphics[width=\textwidth]{introduction/pandemic-timeline}
    \caption[Timeline of pandemics.]{%
        Timeline of pandemics.
        See \url{https://ourworldindata.org/historical-pandemics} for details and sources.
        Licensed under \href{https://creativecommons.org/licenses/by/4.0/}{CC-BY} by Saloni Dattani, Klara Auerbach, Marwa Boukarim, and Max Roser.
    }
    \label{intro:fig:pandemic-timeline}
\end{figure}

Fully quantifying the burden of historical pandemics is challenging due to the absence of systematic records.
However, our best records suggests that, over the last 400 years, they have (on average) occurred every four years with a mean death toll of 5.6 million\todo{cite Marani et al.}.
Modelling of our current situation suggests that we should expect X deaths per year from pandemics\todo{get figures from and cite Madhov et al.}.

\section{Pandemic and epidemic surveillance}

\todo[inline]{Importance of pandemic surveillance, and epidemic v pandemic}

No single quantity summarises the full state of an epidemic, and no single data source fully informs all quantities of interest.
Therefore, a range of metrics and surveillance systems are required to understand the state of an epidemic and the danger of a pathogen (a pathogen is any biological agent that can cause disease).

\subsection{Metrics}

Various forms of \emph{reproduction number} were prominent throughout the COVID-19 pandemic\todo{cite Pellis et al. discussion paper? or maybe something they cite}.
\todo[inline]{Define different reproduction numbers}

\todo[inline]{Talk about growth rates}

Growth rates and reproduction numbers only provide information on how an epidemic is changing.
\emph{Incidence} is the primary quantity of interest for the current (or historical) magnitude of a pandemic.

\todo[inline]{Insert discussion on points in following comment}
% - Why is incidence and prevalence estimation challenging?
% 	- Asymptomatic infection
% 	- Ascertainment bias
% - How do we estimate prevalence
% 	- CIS and REACT
% - Issues with current incidence estimation

\todo[inline]{Define transmission}

To translate any of these quantities into the burden of the disease, a measure of severity is required.
For example, the IFR (infection fatality rate), the proportion of infected individuals who die.
Severity does not feature in this thesis, but it is important to note that it is a key component of understanding the burden of a pandemic.\todo{maybe reword sentence}

\subsection{Data sources}

A wide variety of designed studies, administrative data, and other data sources have been used for pandemic and epidemic surveillance.
In the UK SARS-CoV-2 pandemic, the most prominent were: community testing (pillar 2), the number of severe events (\eg hospital admissions or deaths), seroprevalence testing (seroprevalence is the proportion of people with antibodies to a disease, a marker of prior infection\todo{maybe define antibodies and seroprevalence / make this more succint}), and prevalence surveys (\eg REACT and CIS).\todo{add citations for data sources and spell out REACT and CIS}

Each of these data sources have strengths and weaknesses.
\todo[inline]{compare different data sources and why I focus on CIS}

In addition to population-level surveillance, cohort studies help understand other aspects of the pathogen.
For example, contact tracing data can inform how likely it is for transmission to occur in different settings and/or between different groups of people.
The ATACCC (Assessment of Transmission and Contagiousness of COVID-19 in Contacts) study is an example of a cohort study in the UK SARS-CoV-2 pandemic\todo{add citation}.

@@

\begin{itemize}
    \item Most commonly used types of data: severe events (deaths/hospitalisations), widespread testing (pillar 2), 
    \item Advantages of designed surveys (REACT, CIS)
    \item Use of cohorts (ATACCC)
    \item 
\end{itemize}

\subsection{Estimating incidence and transmission}

\subsubsection{Prior to SARS-CoV-2}

@@

@@

@@

\subsubsection{For SARS-CoV-2}

\todo[inline]{Statistical modelling performed during the pandemic. Discuss broadly but especially focus on those using CIS or REACT. Discuss my contribution in this context.}

\subsection{Estimating duration}

@@

\section{Bayesian inference}

@@

\todo[inline]{Insert contributions listed in Google Doc}

@@

Where possible, aggregated, non-disclosive versions of the dataset are available alongside the code.
The original ATACCC data used is found in the same location.
The original CIS data can only be accessed by accredited researchers within the Office for National Statistic's Secure Research Service\todo{insert details of applying to SRS}.
